A Process for the Emulation of Comparative Oncology Trials with Real-world Evidence (ENCORE)

### 2024-12-15 manuscript/manuscript.qmd
working on plot

format: 
  docx:
    fig-cap-location: top
    tbl-cap-location: top
    reference-doc: custom-reference-doc.docx
  pdf:
    include-in-header: 
      text: |
        \usepackage{lscape}
        \newcommand{\blandscape}{\begin{landscape}}
        \newcommand{\elandscape}{\end{landscape}}
    fig-cap-location: top
    tbl-cap-location: top

@@

library(here)
library(dplyr)
library(knitr)
library(gt)
library(readxl)
library(fs)
library(webshot2)
library(tibble)
library(ggplot2)
library(cowplot)
library(mice)
library(MatchThem)
library(cobalt)
```

@@

To ensure a transparent, reproducible and consistent way of deriving analytic cohorts and performing statistical analyses across trial emulations, we developed and internal R package `encore.io`with parameterized functions. A detailed documentation can be found in the Supplementary Material.

@@

-   **Estimate agreement**: agreement that the estimated RWE treatment treatment effect is within the 95% CI of the RCT treatment effect estimate. Provided that for some emulations the power of the RWE study may be larger than that of the original RCT, this could lead to situations where there us no regulatory agreement although the treatment effect estimates are highly overlapping but the RCT estimate crosses the null (or vice versa).

@@

We would like to note that only the SMD agreement metric will be applicable to absolute risk measures like median survival times or survival probabilities.

@@

Target populations in international trials...genetic setup.

@@

# save
suppressMessages(ggsave(
  plot = plot_balance,
  filename = "Figure_3_balance.png",
  path = here("manuscript"),
  dpi = 800,
  width = 15,
  height = 10
  ))

@@

>>>>

format: 
  docx:
    fig-cap-location: top
    tbl-cap-location: top
    reference-doc: custom-reference-doc.docx
  pdf:
    fig-cap-location: top
    tbl-cap-location: top

@@

library(here)
library(dplyr)
library(knitr)
library(gt)
library(readxl)
library(fs)
library(webshot2)
library(tibble)
library(ggplot2)
library(cowplot)
library(mice)
library(MatchThem)
library(cobalt)
library(gtExtras)
```

@@

To ensure a transparent, reproducible and consistent way of deriving analytic cohorts and performing statistical analyses across trial emulations, we developed and internal R package `encore.io` with parameterized functions. A detailed documentation can be found in the Supplementary Material.

@@

-   **Estimate agreement**: agreement that the estimated RWE treatment treatment effect is within the 95% CI of the RCT treatment effect estimate. Provided that for some emulations, the power of the RWE study may be larger than that of the original RCT, this could lead to situations where there is no regulatory agreement although the treatment effect estimates are highly overlapping but with the RCT estimate crossing the null (or vice versa in case the RCT has a larger power than the RWE emulation).

@@

For the secondary endpoints of interest (e.g., median survival time or survival probabilities) only the SMD agreement metric will be applicable.

@@

Building off an established approach based on the RCT DUPLICATE project[@wang2023emulation; @franklin2020nonrandomized], the ENCORE project aims to emulate 12 oncology trials using multiple EHR data sources to assess the potential use of RWE for regulatory science. The project focuses on four common cancers and will evaluate the agreement of treatment effect estimates between RCTs and their respective emulations.

-   Distributions of target populations in international trials...genetic setup.

-   Bias towards the null, per protocol effect difficult

-   Transportability =\> RWE estimates may not be the same as RCT but still informative given more generalizable/broader populations considered

@@

# save
suppressMessages(ggsave(
  plot = plot_balance,
  filename = "Figure_3_balance.png",
  path = here("manuscript"),
  dpi = 800,
  width = 15,
  height = 8
  ))

@@

```{r}
#| label: fig-metrics
#| fig-cap: "Visualization of agreement metrics example."
#| fig-widh: 4
#| fig-height: 7

source(here("functions","smd_agreement.R"))

data <- tribble(
  ~group, ~estimate, ~lower, ~upper,
  "RCT", 0.75, 0.61, 0.91,
  "RWE", 0.80, 0.51, 1.1
  )

smd <- smd_agreement(
  rct_estimate = data |> filter(group == "RCT") |> pull(estimate),
  rct_lower = data |> filter(group == "RCT") |> pull(lower),
  rct_upper = data |> filter(group == "RCT") |> pull(upper),
  rwe_estimate = data |> filter(group == "RWE") |> pull(estimate),
  rwe_lower = data |> filter(group == "RWE") |> pull(lower),
  rwe_upper = data |> filter(group == "RWE") |> pull(upper)
  )

data |> 
  gt() |> 
  gt_plt_conf_int(
    column = estimate, 
    ci_columns = c(lower, upper),
    ref_line = 1,
    palette = c("red", "lightgrey", "black", "red")
    ) |> 
  cols_hide(c(lower, upper)) |> 
  # add agreement metrics
  cols_add(
    metric = c(md("Regulatory agreement: <br><br> Estimate agreement:"), "SMD:"),
    results = c(md("No <br><br> Yes"), ifelse(abs(smd) < 1.96, "Yes", "No"))
    ) |> 
  tab_options(
    table_body.hlines.width = px(0)
  ) |> 
  cols_label(
    group = "",
    estimate = "HR (95% CI)",
    metric = "Agreement metrics",
    results = ""
    ) |> 
  fmt_markdown() |> 
  tab_style(
    style = cell_text(align = "center", v_align = "middle"),
    locations = cells_column_labels()
    ) |> 
  tab_style(
    style = cell_text(weight = "bold"),
    locations = cells_column_labels()
    )
```
